### 2014-01-31 platforms.tex
platforms: android

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{hyperref}
\RequirePackage[date=terse,isbn=true,doi=true,url=false,maxbibnames=9,backref=true]{biblatex}
\addbibresource{securityupdates.bib}
\renewcommand{\bibfont}{\small}

\AtEveryBibitem{% Clean up the bibtex rather than editing it
 \clearname{editor} % remove editors
}

\author{Daniel Thomas}
\title{Platforms for security update research}
\begin{document}
\maketitle

\section{Introduction}
Early computers allowed running code to do anything, later more restrictive systems were developed which had an operating system which could do anything but which ran programs which were restricted in what they could do.
However mostly this was broken down on a per user basis, code running as a particular user could do anything a user could do.
More recently there have been efforts at greater compartmentalisation and more fine grained permissions for applications.
For example the Android permission system restricts what each app can do in a more precise way than installing applications on Linux does.
This fine grained compartmentalisation only works if there are no known vulnerabilities which break it and allow code to do arbitrary things.
Unfortunately this is not the case.

\section{Android}
In Android each app runs in a separate user account and is granted certain permissions statically after user review on installation.
Apps can use the rich Android API and embed native code which can make use of the Linux API.
The security boundary is the user but there have been many examples where it is possible for an app to exploit a vulnerability in Android and gain root permissions and do anything.
To analyse the proportion of the time deployed Android devices are exposed to vulnerabilities which break their security boundary we can use the Device Analyzer data to look at running versions and the \url{http://androidvulnerabilities.org} website to find out which vulnerabilities applied to which versions at what time.

\section{iOS}
\section{Firefox}
\section{Chrome}
\section{ChromeOS}
\section{JVM}
\section{Xen}
\section{VMWare}

\printbibliography

\end{document}