### 2014-11-13 avostats.tex
Updated avostats with larger vulnerabilities table

\newcommand{\avoNumSubmitters}{7}
\newcommand{\avoTotalExternalLines}{25 Million}
\newcommand{\avoNumExternalProjects}{176}
\newcommand{\avoNumBigExternalLinesOfCode}{25 Million}
\newcommand{\avoBigExternalLinesOfCodePerc}{99.7\%}
\newcommand{\avoNumVulnerabilities}{31}
\newcommand{\avoNumVulnAllAndroid}{15}
\newcommand{\avoNumVulnSpecific}{16}
\newcommand{\avoStartDate}{2013-08-28}
\newcommand{\avoEndDate}{2014-10-17}
\newcommand{\avoFirstDataDate}{2010-07-13}
\newcommand{\avoLastDataDate}{2014-07-29}
\newcommand{\avoVulnsPerYearNominal}{7.66}
\newcommand{\avoVulnsPerYear}{$7.66 \pm 1.38$}
\newcommand{\avoVulnsPerYearAllAndroidNominal}{3.71}
\newcommand{\avoVulnsPerYearAllAndroid}{$3.71 \pm 0.957$}
\newcommand{\avoTabAndVulns}{\begin{table} \centering \small \begin{tabular}{l|l|c} Vulnerability & How known & Date \\ \hline KillingInTheNameOf & Fixed on & 2010-07-13 \\ exploid udev & Discovered on & 2010-07-15 \\ levitator & Discovered on & 2011-03-10 \\ Gingerbreak & Fixed on & 2011-04-18 \\ zergRush & Discovered on & 2011-10-06 \\ APK duplicate file & Discovered on & 2013-02-18 \\ APK unchecked name & Discovered on & 2013-06-30 \\ APK unsigned shorts & Fixed on & 2013-07-03 \\ vold asec & Fixed on & 2014-01-27 \\ Fake ID & Fixed on & 2014-04-17 \\ TowelRoot & Discovered on & 2014-05-03 \\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
\newcommand{\avoNumBigExternalProjects}{40}
\newcommand{\avoNumAnalysedExternalProjects}{28}
\newcommand{\avoNumAnalysedExternalLinesOfCode}{6 Million}
\newcommand{\avoAnalysedExternalLinesOfCodePerc}{24.9\%}
\newcommand{\avoBigExternalMedianVersions}{1.5}
\newcommand{\avoBigExternalMeanVersionsNominal}{2.07}
\newcommand{\avoBigExternalMeanVersions}{$2.07 \pm 1.44$}
\newcommand{\avoBigExternalTotalVersions}{58}