### 2015-01-16 rebuttal.txt
Updated rebuttal

We don't claim to make a technical contribution.
The call is for "advances in the theory, design, implementation, analysis, verification, or empirical evaluation and measurement of secure systems."
We have provided an empirical evaluation and measurement of Android.

We have confirmed the conventional wisdom, but provide solid evidence.
In this context, this paper is the solution to the problem of slow Android updates: public data comparing manufacturers and device models, provides market pressure for industry improvements in security.
We will publish (and update) this information on androidvulnerabilities.org.

Regarding the Bangladesh study anomaly:
Truncating our data before it starts increases the percentage of vulnerable devices from 87.6% to 87.7% and decreases the percentage of devices running the latest version from 5.58% to 5.43%.

When we see a new version on one device, often we see all other devices with the same model receive the update promptly, regardless of operator or user behaviour.
By looking at our data, 50% of devices update to a newly available build number within 26 days, with a long tail thereafter.
Update behaviour is consistent within device models with some being fast and others slow to deploy.
This means that the user is unlikely to be the problem and that the bottleneck is with Google, the Manufacturer or, for some devices, the Operator.
This is based on the significant devices shown in Table V.

This means that the bias in the data doesn't affect our main conclusions
about particular manufacturers or device models (i.e. "buy a Nexus device").
While bias will affect which device models we analyse it will not affect the result of that analysis.

Short contributions are not a problem as we are getting a sample from the distribution of all the devices.

We have zero cases where Device Analyzer has been broken by an OS update, Android has good backwards compatibility.

"On the other hand, when the proportion of vulnerable devices drops,
it might be due to vulnerable devices leaving, or patched devices
joining." -> If a user stops using their phone because it is vulnerable,
then this is okay: as it is no longer in circulation. It is unlikely that a
user would discover their phone is vulnerable, and uninstall DA but carry
on using it.

We the paper explains the reason for the exp in the scoring function just before Equation 1.

We can of course construct any scoring function we like! The paper explains
why the three variables are useful. A linear combination of
f, u and 1/exp(m) is reasonable because we are interested in devices which
perform well against all three measures.

We can expand Table II if useful.