### 2015-04-24 wisec2015abstract.tex
Cut out irrlevant stuff and update other content in wisec abstract

\documentclass{sig-alternate}
\pagenumbering{arabic}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}

\RequirePackage[date=terse, isbn=true, doi=true, url=false, urldate=iso8601, maxbibnames=9, backref=false, backend=bibtex, style=ieee]{biblatex}
\addbibresource{securityupdates.bib}
\renewcommand{\bibfont}{\small}

\AtEveryBibitem{% Clean up the bibtex rather than editing it
 \clearname{editor} % remove editors
}

\include{dastats}
\newcommand{\da}{Device Analyzer}

\numberofauthors{4}
\author{Daniel~R.~Thomas \and Daniel~T.~Wagner \and Alastair~R.~Beresford \and Andrew~Rice\\       \email{Firstname.Lastname@cl.cam.ac.uk}
}
\title{POSTER: \da :\\Measuring the security of smartphones}
\conferenceinfo{WiSec}{2015 New York City, USA}
\begin{document}
\maketitle

TODO

Our Device Analyzer application aims to collect a large-scale research data-set of phone usage.
Device Analyzer gathers data about running processes, wireless connectivity, the phone’s location, GSM communication, battery state and a number of system parameters.
We want to extend the range of data collected as well as the breadth of the targeted population.
Today’s smart phones are able to capture much richer data than was available some years ago while at the same time the Android Market allows us to distribute the app to a broader audience than was possible before.

\section{Data Collection}
Device Analyzer makes use of events from the Android operating system to record many changes in system state and polls for information not provided by events.
All data are stored locally in a timestamped key-value log file for periodic upload to a central server.
By default Device Analyzer attempts to minimise the impact of this by scheduling uploads when the device is charging and has WiFi.

\section{Timestamps}
Simple event timestamping using the current system time is inadequate in a number of cases.
System time is subject to discontinuities, such as when the user manually changes the time on their phone or when the cell network broadcasts a correction.
Furthermore, local time is not always available: we see that for some seconds after a system upgrade the returned dates are in the 1980s!
To provide dependable timestamps in the face of such disruptions, we use the system uptime as our reference frame as it is supposed to be guaranteed to not jump backwards or forwards.
Changes to the mapping of uptime to local time are logged in the event stream and post processed to correct for errors~\cite{Wagner2013}.

\section{Viewing Data}
Data can be viewed on the phone itself and on the web after it was uploaded.
The data displayed on the phone is designed to show the user which kinds of data are collected and provides summary graphs.

\section{Hashing and Aliases}
Rather than uploading identifying user data such as phone numbers or network addresses in plain text, they are hashed with a device-specific salt to maintain user privacy.
Users can choose to assign private nicknames to these anonymous values so that they can identify them when they log in to the website.
The dataset itself contains only the hashed values.

\section{Privacy Concerns}
Have released release the collected anonymous dataset where participants have given us permission to do so to researchers at many organisations after their agreeing to terms which protect the privacy of the users.
We only publish data that was collected at least three months ago, giving participants time to view the data they contributed.
If they are concerned about having their data released they can opt-out of any further data collection and erase all unpublished data from our server.
Furthermore, while physical location of participants could be an asset of our dataset, such data needs to be handled with great care, as disclosure of even approximate home/work location pairs could lead to deanonymization of participants~\cite{Golle2009}.
Hence we only provide the country the device was most often in to the researchers.

\section{Conclusion}
A rich dataset of smart phone usage for a large population has undeniable uses for device owners as well as providers and manufacturers.
However, collecting such detailed data raises privacy concerns even if the data are never uploaded if an adversary gains physical access to the device.
We have tried to address privacy issues which arise so that informed consent can be assumed.
Please do contact us if you wish to participate either by contributing data or conducting research using it.

\section{Acknowledgements}
This work was supported by the University of Cambridge Computer Laboratory Premium Studentship scheme, a Google focussed research award and the EPSRC Standard Research Grant EP/P505445/1.

\printbibliography

\end{document}