### 2015-07-27 avostats.tex
Tweak some tables and hyphenation rules to make things fit better inside the columns

\newcommand{\avoNumSubmitters}{7}
\newcommand{\avoTotalExternalLines}{25 Million}
\newcommand{\avoNumExternalProjects}{176}
\newcommand{\avoNumBigExternalLinesOfCode}{25 Million}
\newcommand{\avoBigExternalLinesOfCodePerc}{99.7\%}
\newcommand{\avoNumVulnerabilities}{32}
\newcommand{\avoNumVulnAllAndroid}{15}
\newcommand{\avoNumVulnSpecific}{17}
\newcommand{\avoStartDate}{2013-08-28}
\newcommand{\avoEndDate}{2015-03-24}
\newcommand{\avoFirstDataDateMonth}{July 2010}
\newcommand{\avoFirstDataDate}{2010-07-13}
\newcommand{\avoLastDataDateMonth}{March 2015}
\newcommand{\avoLastDataDate}{2015-03-12}
\newcommand{\avoVulnsPerYearNominal}{6.86}
\newcommand{\avoVulnsPerYear}{$6.86 \pm 1.21$}
\newcommand{\avoVulnsPerYearAllAndroidNominal}{3.21}
\newcommand{\avoVulnsPerYearAllAndroid}{$3.21 \pm 0.83$}
\newcommand{\avoVulnsPerYearTwosfNominal}{6.9}
\newcommand{\avoVulnsPerYearTwosf}{$6.9 \pm 1.2$}
\newcommand{\avoVulnsPerYearAllAndroidTwosfNominal}{3.2}
\newcommand{\avoVulnsPerYearAllAndroidTwosf}{$3.2 \pm 0.0$}
\newcommand{\avoTabAndVulns}{\begin{table} \centering \small \begin{tabular}{l|l|c|c} Vulnerability & How known & Date & Categories\\ \hline KillingInTheNameOf & Fixed & 2010-07-13 & sys, kern\\ exploid udev & Discovered & 2010-07-15 & kernel\\ levitator & Discovered & 2011-03-10 & kernel\\ Gingerbreak & Fixed & 2011-04-18 & system\\ zergRush & Discovered & 2011-10-06 & system\\ APK duplicate file & Discovered & 2013-02-18 & signature\\ APK unchecked name & Discovered & 2013-06-30 & signature\\ APK unsigned shorts & Fixed & 2013-07-03 & signature\\ vold asec & Fixed & 2014-01-27 & system\\ Fake ID & Fixed & 2014-04-17 & signature\\ TowelRoot & Discovered & 2014-05-03 & kernel\\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
\newcommand{\avoNumBigExternalProjects}{40}
\newcommand{\avoNumAnalysedExternalProjects}{28}
\newcommand{\avoNumAnalysedExternalLinesOfCode}{6 Million}
\newcommand{\avoAnalysedExternalLinesOfCodePerc}{24.9\%}
\newcommand{\avoBigExternalMedianVersions}{2.0}
\newcommand{\avoBigExternalMeanVersionsNominal}{2.57}
\newcommand{\avoBigExternalMeanVersions}{$2.57 \pm 1.84$}
\newcommand{\avoBigExternalTotalVersions}{72}