### 2015-10-02 avostats.tex
./update.sh

\newcommand{\avoNumSubmitters}{8}
\newcommand{\avoTotalExternalLines}{25 million}
\newcommand{\avoNumExternalProjects}{176}
\newcommand{\avoNumBigExternalLinesOfCode}{25 million}
\newcommand{\avoBigExternalLinesOfCodePerc}{99.7\%}
\newcommand{\avoNumVulnerabilities}{33}
\newcommand{\avoNumVulnAllAndroid}{16}
\newcommand{\avoNumVulnSpecific}{17}
\newcommand{\avoStartDate}{2013-08-28}
\newcommand{\avoEndDate}{2015-03-24}
\newcommand{\avoFirstDataDateMonth}{July 2010}
\newcommand{\avoFirstDataDate}{2010-07-13}
\newcommand{\avoLastDataDateMonth}{March 2015}
\newcommand{\avoLastDataDate}{2015-03-12}
\newcommand{\avoVulnsPerYearNominal}{7.07}
\newcommand{\avoVulnsPerYear}{$7.07 \pm 1.23$}
\newcommand{\avoVulnsPerYearAllAndroidNominal}{3.43}
\newcommand{\avoVulnsPerYearAllAndroid}{$3.43 \pm 0.85$}
\newcommand{\avoVulnsPerYearTwosfNominal}{7.1}
\newcommand{\avoVulnsPerYearTwosf}{$7.1 \pm 1.2$}
\newcommand{\avoVulnsPerYearAllAndroidTwosfNominal}{3.4}
\newcommand{\avoVulnsPerYearAllAndroidTwosf}{$3.4 \pm 0.0$}
\newcommand{\avoTabAndVulns}{\begin{table} \centering \small \begin{tabular}{l|l|c|c} Vulnerability & How known & Date & Categories\\ \hline KillingInTheNameOf & Fixed on & 2010-07-13 & system, kernel\\ exploid udev & Discovered on & 2010-07-15 & kernel\\ levitator & Discovered on & 2011-03-10 & kernel\\ Gingerbreak & Fixed on & 2011-04-18 & system\\ zergRush & Discovered on & 2011-10-06 & system\\ APK duplicate file & Discovered on & 2013-02-18 & signature\\ APK unchecked name & Discovered on & 2013-06-30 & signature\\ APK unsigned shorts & Fixed on & 2013-07-03 & signature\\ vold asec & Fixed on & 2014-01-27 & system\\ Fake ID & Fixed on & 2014-04-17 & signature\\ TowelRoot & Discovered on & 2014-05-03 & kernel\\ ObjectInputStream_deserializable & Discovered on & 2014-06-22 & system\\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
\newcommand{\avoNumBigExternalProjects}{40}
\newcommand{\avoNumAnalysedExternalProjects}{28}
\newcommand{\avoNumAnalysedExternalLinesOfCode}{6 million}
\newcommand{\avoAnalysedExternalLinesOfCodePerc}{24.9\%}
\newcommand{\avoBigExternalMedianVersions}{2.0}
\newcommand{\avoBigExternalMeanVersionsNominal}{2.57}
\newcommand{\avoBigExternalMeanVersions}{$2.57 \pm 1.84$}
\newcommand{\avoBigExternalTotalVersions}{72}