### 2015-10-02 dastats.tex

\newcommand{\daNumSigModels}{18}
\newcommand{\daSigNumDevices}{100}
\newcommand{\daSigNumDeviceDays}{10\,000}
\newcommand{\daSigNumDays}{100}
\newcommand{\daSigNumDevicesDay}{20}
\newcommand{\daSigVersionPerc}{1\%}
\newcommand{\daSigVersionDays}{10}
\newcommand{\daNumVulnsUsed}{11}
\newcommand{\daNumDeviceDataDevices}{50}
\newcommand{\daOSYearsOfData}{4}
\newcommand{\daOSMonthsOfData}{51}
\newcommand{\daStartDate}{2011-07-01}
\newcommand{\daEndDate}{2015-09-22}
\newcommand{\daFullDeployedAt}{95\%}
\newcommand{\daNumManufacturers}{302}
\newcommand{\daNumSigManufacturers}{10}
\newcommand{\daNumModels}{2\,770}
\newcommand{\daNumOperators}{1\,470}
\newcommand{\daNumSigOperators}{14}
\newcommand{\daNumOSDevices}{20\,500}
\newcommand{\daOSVersionPercValidLines}{99.9\%}
\newcommand{\daOSTotalDaysData}{1\,350\,000}
\newcommand{\daMeanInsecurityPercNominal}{87.3\%}
\newcommand{\daMeanInsecurityPerc}{$87.3 \pm 0.0\%$}
\newcommand{\daMeanInsecurityPercTwosfNominal}{87\%}
\newcommand{\daMeanInsecurityPercTwosf}{$87 \pm 0\%$}
\newcommand{\daMeanSecurityPercNominal}{12.7\%}
\newcommand{\daMeanSecurityPerc}{$12.7 \pm 0.0\%$}
\newcommand{\daMeanSecurityPercTwosfNominal}{13\%}
\newcommand{\daMeanSecurityPercTwosf}{$13 \pm 0\%$}
\newcommand{\daVulnFreeNominal}{0.127}
\newcommand{\daVulnFree}{$0.127 \pm 0.0$}
\newcommand{\daMeanOutstandingVulnerabilities}{$0.509 \pm 0.0$}
\newcommand{\daUpdatednessNominal}{0.0553}
\newcommand{\daUpdatedness}{$0.0553 \pm 0.0$}
\newcommand{\daUpdatednessPercNominal}{5.53\%}
\newcommand{\daUpdatednessPerc}{$5.53 \pm 0.0\%$}
\newcommand{\daUpdatednessPercTwosfNominal}{5.5\%}
\newcommand{\daUpdatednessPercTwosf}{$5.5 \pm 0.0\%$}
\newcommand{\daSecurityScoreNominal}{2.93}
\newcommand{\daSecurityScore}{$2.93 \pm 0.0$}
\newcommand{\daCurrentMeanInsecurityPercNominal}{68.8\%}
\newcommand{\daCurrentMeanInsecurityPerc}{$68.8 \pm 0.0\%$}
\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{69\%}
\newcommand{\daCurrentMeanInsecurityPercTwosf}{$69 \pm 0\%$}
\newcommand{\daCurrentMeanSecurityPercNominal}{31.2\%}
\newcommand{\daCurrentMeanSecurityPerc}{$31.2 \pm 0.2\%$}
\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{31\%}
\newcommand{\daCurrentMeanSecurityPercTwosf}{$31 \pm 0\%$}
\newcommand{\daCurrentVulnFreeNominal}{0.312}
\newcommand{\daCurrentVulnFree}{$0.312 \pm 0.002$}
\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.0645 \pm 0.0$}
\newcommand{\daCurrentUpdatednessNominal}{0.0934}
\newcommand{\daCurrentUpdatedness}{$0.0934 \pm 0.0019$}
\newcommand{\daCurrentUpdatednessPercNominal}{9.34\%}
\newcommand{\daCurrentUpdatednessPerc}{$9.34 \pm 0.19\%$}
\newcommand{\daCurrentUpdatednessPercTwosfNominal}{9.3\%}
\newcommand{\daCurrentUpdatednessPercTwosf}{$9.3 \pm 0.0\%$}
\newcommand{\daCurrentSecurityScoreNominal}{4.43}
\newcommand{\daCurrentSecurityScore}{$4.43 \pm 0.0$}
\newcommand{\daDailyParticipationNominal}{872}
\newcommand{\daDailyParticipation}{$872 \pm 271$}
\newcommand{\daNumFullVersions}{1\,300}
\newcommand{\daNumSigFullVersions}{101}
\newcommand{\daNumOSVersions}{50}
\newcommand{\daNumSigOSVersions}{28}
\newcommand{\daNumAPIVersions}{16}
\newcommand{\daNumFullVersionUpdates}{4\,670}
\newcommand{\daNumOSVersionUpdates}{3\,940}
\newcommand{\daUpdatesPerYearNominal}{1.27}
\newcommand{\daUpdatesPerYear}{$1.27 \pm 0.01$}
\newcommand{\daUpdatesPerYearTwosfNominal}{1.3}
\newcommand{\daUpdatesPerYearTwosf}{$1.3 \pm 0.0$}
\newcommand{\daNumFullOnlyVersionUpdates}{724}
\newcommand{\daNumSecurityUpdates}{1\,270}
\newcommand{\daNumPossibleSecurityUpdates}{2\,310}
\newcommand{\daOSCurveFitParamFirst}{83.6}
\newcommand{\daOSCurveFitParamSecond}{\num{0.0028}}
\newcommand{\daOSCurveFitRMSE}{0.114}
\newcommand{\daOSCurvePolyRMSE}{0.114}
\newcommand{\daOSCurveSplineRMSE}{0.114}
\newcommand{\daOSCurveHalfDeployedYearsNominal}{0.908}
\newcommand{\daOSCurveHalfDeployedYears}{$0.908 \pm 0.254$}
\newcommand{\daOSCurveHalfDeployedNominal}{332}
\newcommand{\daOSCurveHalfDeployed}{$332 \pm 92$}
\newcommand{\daOSCurveFullDeployedYearsNominal}{3.16}
\newcommand{\daOSCurveFullDeployedYears}{$3.16 \pm 72$}
\newcommand{\daOSCurveFullDeployedNominal}{1\,150}
\newcommand{\daOSCurveFullDeployed}{$1\,150 \pm 265\,000$}
\newcommand{\daAPICurveFitParamFirst}{97.8}
\newcommand{\daAPICurveFitParamSecond}{\num{0.0028}}
\newcommand{\daAPICurveFitRMSE}{0.11}
\newcommand{\daAPICurvePolyRMSE}{0.111}
\newcommand{\daAPICurveSplineRMSE}{0.111}
\newcommand{\daAPICurveHalfDeployedYearsNominal}{0.946}
\newcommand{\daAPICurveHalfDeployedYears}{$0.946 \pm 0.244$}
\newcommand{\daAPICurveHalfDeployedNominal}{345}
\newcommand{\daAPICurveHalfDeployed}{$345 \pm 89$}
\newcommand{\daAPICurveFullDeployedYearsNominal}{3.2}
\newcommand{\daAPICurveFullDeployedYears}{$3.2 \pm 7$}
\newcommand{\daAPICurveFullDeployedNominal}{1\,170}
\newcommand{\daAPICurveFullDeployed}{$1\,170 \pm 265\,000$}
\newcommand{\daVulnAPKDuplicateFileOctoberPerc}{91.7\%}
\newcommand{\daVulnZergRushMonthsDefFixDeployed}{27.3}
\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.24 \pm 0.00$ & $0.34 \pm 0.00$ & $0.60 \pm 0.01$ & $4.11 \pm 0.02$ \\  Motorola & $0.19 \pm 0.00$ & $0.12 \pm 0.00$ & $0.71 \pm 0.02$ & $3.11 \pm 0.02$ \\  Samsung & $0.13 \pm 0.00$ & $0.04 \pm 0.00$ & $0.58 \pm 0.00$ & $2.80 \pm 0.00$ \\  Sony & $0.14 \pm 0.00$ & $0.19 \pm 0.00$ & $1.09 \pm 0.02$ & $2.66 \pm 0.02$ \\  HTC & $0.14 \pm 0.00$ & $0.10 \pm 0.00$ & $0.86 \pm 0.01$ & $2.65 \pm 0.02$ \\  asus & $0.21 \pm 0.00$ & $0.51 \pm 0.01$ & $6.01 \pm 0.07$ & $2.38 \pm 0.02$ \\  other & $0.06 \pm 0.00$ & $0.05 \pm 0.00$ & $1.00 \pm 0.01$ & $2.02 \pm 0.02$ \\  alps & $0.03 \pm 0.00$ & $0.19 \pm 0.01$ & $3.99 \pm 0.08$ & $0.80 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $5.00 \pm 0.05$ & $0.32 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $6.00 \pm 0.08$ & $0.27 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
\newcommand{\daSecScoreBestmanufacturer}{LG}
\newcommand{\daSecScoreBestmanufacturerScoreNominal}{4.11}
\newcommand{\daSecScoreBestmanufacturerScore}{$4.11 \pm 0.0$}
\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{142}
\newcommand{\daSecScoreWorstmanufacturer}{walton}
\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.273}
\newcommand{\daSecScoreWorstmanufacturerScore}{$0.273 \pm 0.007$}
\newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{21}
\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Galaxy Nexus & $0.50 \pm 0.00$ & $0.54 \pm 0.01$ & $1.53 \pm 0.04$ & $4.70 \pm 0.04$ \\  Nexus 4 & $0.33 \pm 0.00$ & $0.82 \pm 0.01$ & $6.06 \pm 0.09$ & $3.78 \pm 0.04$ \\  Nexus 7 & $0.27 \pm 0.00$ & $0.73 \pm 0.01$ & $5.92 \pm 0.09$ & $3.29 \pm 0.04$ \\  other & $0.10 \pm 0.00$ & $0.14 \pm 0.00$ & $0.51 \pm 0.00$ & $3.08 \pm 0.00$ \\  Desire HD & $0.08 \pm 0.00$ & $0.05 \pm 0.00$ & $0.38 \pm 0.02$ & $2.91 \pm 0.04$ \\  HTC Sensation & $0.35 \pm 0.00$ & $0.01 \pm 0.01$ & $1.57 \pm 0.05$ & $2.44 \pm 0.05$ \\  GT-I9100 & $0.22 \pm 0.00$ & $0.02 \pm 0.00$ & $1.23 \pm 0.02$ & $2.27 \pm 0.02$ \\  HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.06$ & $1.74 \pm 0.07$ \\  GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.52 \pm 0.05$ & $1.43 \pm 0.02$ \\  GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.79 \pm 0.06$ & $0.90 \pm 0.05$ \\  GT-I9505 & $0.06 \pm 0.00$ & $0.13 \pm 0.00$ & $6.82 \pm 0.07$ & $0.62 \pm 0.01$ \\  GT-I9300 & $0.13 \pm 0.00$ & $0.01 \pm 0.00$ & $6.23 \pm 0.04$ & $0.57 \pm 0.01$ \\  HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.03 \pm 0.05$ & $0.28 \pm 0.03$ \\  GT-N7100 & $0.06 \pm 0.00$ & $0.00 \pm 0.01$ & $6.93 \pm 0.08$ & $0.24 \pm 0.02$ \\  Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.12$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
\newcommand{\daSecScoreBestmodel}{Galaxy Nexus}
\newcommand{\daSecScoreBestmodelScoreNominal}{4.7}
\newcommand{\daSecScoreBestmodelScore}{$4.7 \pm 0.0$}
\newcommand{\daSecScoreBestmodelNumFullVersions}{49}
\newcommand{\daSecScoreWorstmodel}{Symphony W68}
\newcommand{\daSecScoreWorstmodelScoreNominal}{0.0001}
\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0272$}
\newcommand{\daSecScoreWorstmodelNumFullVersions}{1}
\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline nexus & $0.40 \pm 0.00$ & $0.49 \pm 0.00$ & $0.55 \pm 0.01$ & $5.28 \pm 0.02$ \\  non-nexus & $0.11 \pm 0.00$ & $0.03 \pm 0.00$ & $0.51 \pm 0.00$ & $2.76 \pm 0.00$ \\ \end{tabular} \caption{Security scores for nexus} \label{tab:sec_summary} \end{table*}}
\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.28 \pm 0.00$ & $0.12 \pm 0.00$ & $0.37 \pm 0.02$ & $3.91 \pm 0.03$ \\  T-Mobile & $0.22 \pm 0.00$ & $0.19 \pm 0.00$ & $0.39 \pm 0.01$ & $3.87 \pm 0.02$ \\  Orange & $0.22 \pm 0.00$ & $0.10 \pm 0.00$ & $0.36 \pm 0.02$ & $3.64 \pm 0.04$ \\  3 & $0.22 \pm 0.00$ & $0.09 \pm 0.00$ & $0.47 \pm 0.02$ & $3.45 \pm 0.03$ \\  Sprint & $0.18 \pm 0.00$ & $0.11 \pm 0.00$ & $0.43 \pm 0.02$ & $3.41 \pm 0.03$ \\  Vodafone uk & $0.14 \pm 0.00$ & $0.13 \pm 0.00$ & $0.52 \pm 0.03$ & $3.19 \pm 0.04$ \\  AT\&T & $0.15 \pm 0.00$ & $0.08 \pm 0.00$ & $0.43 \pm 0.02$ & $3.17 \pm 0.02$ \\  unknown & $0.12 \pm 0.00$ & $0.20 \pm 0.00$ & $0.80 \pm 0.01$ & $2.95 \pm 0.02$ \\  Verizon & $0.19 \pm 0.00$ & $0.09 \pm 0.00$ & $0.82 \pm 0.02$ & $2.86 \pm 0.02$ \\  n Telenor & $0.05 \pm 0.00$ & $0.12 \pm 0.00$ & $1.14 \pm 0.02$ & $2.01 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.03 \pm 0.00$ & $1.44 \pm 0.03$ & $1.45 \pm 0.03$ \\  Grameenphone & $0.01 \pm 0.00$ & $0.04 \pm 0.00$ & $1.66 \pm 0.02$ & $1.10 \pm 0.01$ \\  Robi & $0.00 \pm 0.00$ & $0.08 \pm 0.00$ & $2.08 \pm 0.04$ & $0.91 \pm 0.03$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $2.55 \pm 0.04$ & $0.55 \pm 0.02$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
\newcommand{\daSecScoreBestoperator}{O2 uk}
\newcommand{\daSecScoreBestoperatorScoreNominal}{3.91}
\newcommand{\daSecScoreBestoperatorScore}{$3.91 \pm 0.0$}
\newcommand{\daSecScoreBestoperatorNumFullVersions}{103}
\newcommand{\daSecScoreWorstoperator}{banglalink}
\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.547}
\newcommand{\daSecScoreWorstoperatorScore}{$0.547 \pm 0.017$}
\newcommand{\daSecScoreWorstoperatorNumFullVersions}{75}
\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ \hline equal & 0.0 & 0.0 & 0.0714 & 0.0 \\weight $m$ & 0.2 & 0.333 & 0.214 & 0.0 \\weight $u$ & 0.2 & 0.0 & 0.143 & 0.0 \\$f$ & 0.3 & 0.733 & 0.357 & 0.0 \\$m$ & 0.6 & 0.733 & 0.357 & 0.5 \\$u$ & 0.9 & 0.667 & 0.786 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
\newcommand{\daTabChangeInScores}{\begin{table*}\small \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ \hline $m$ & $-1.67 \pm 2.26$ & $-0.65 \pm 2.56$ & $-3.34 \pm 1.27$ & $-3.4 \pm 1.9$ \\weight $m$ & $-0.263 \pm 0.278$ & $-0.0958 \pm 0.325$ & $-0.464 \pm 0.159$ & $-0.488 \pm 0.225$ \\equal & $-0.106 \pm 0.061$ & $-0.0343 \pm 0.111$ & $-0.145 \pm 0.043$ & $-0.164 \pm 0.035$ \\weight $u$ & $-0.0569 \pm 0.121$ & $-0.00706 \pm 0.225$ & $0.029 \pm 0.082$ & $-0.00391 \pm 0.12$ \\$u$ & $0.389 \pm 1.62$ & $0.238 \pm 2.09$ & $1.6 \pm 0.9$ & $1.44 \pm 1.51$ \\$f$ & $0.958 \pm 0.556$ & $0.309 \pm 1.0$ & $1.31 \pm 0.39$ & $1.48 \pm 0.32$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
\newcommand{\daGPAPICurveFitParamFirst}{82.1}
\newcommand{\daGPAPICurveFitParamSecond}{\num{0.00257}}
\newcommand{\daGPAPICurveFitRMSE}{0.176}
\newcommand{\daGPAPICurvePolyRMSE}{0.176}
\newcommand{\daGPAPICurveSplineRMSE}{0.176}
\newcommand{\daGPAPICurveHalfDeployedYearsNominal}{0.964}
\newcommand{\daGPAPICurveHalfDeployedYears}{$0.964 \pm 0.462$}
\newcommand{\daGPAPICurveHalfDeployedNominal}{352}
\newcommand{\daGPAPICurveHalfDeployed}{$352 \pm 169$}
\newcommand{\daGPAPICurveFullDeployedYearsNominal}{3.42}
\newcommand{\daGPAPICurveFullDeployedYears}{$3.42 \pm 79$}
\newcommand{\daGPAPICurveFullDeployedNominal}{1\,250}
\newcommand{\daGPAPICurveFullDeployed}{$1\,250 \pm 289\,000$}
\newcommand{\daGPAPIPerAPIMeanNominal}{0.00472}
\newcommand{\daGPAPIPerAPIMean}{$0.00472 \pm 0.00327$}
\newcommand{\daGPAPIPerAPIStdevNominal}{0.0209}
\newcommand{\daGPAPIPerAPIStdev}{$0.0209 \pm 0.0121$}
\newcommand{\daGPAPIPerAPIAbsMeanNominal}{0.0143}
\newcommand{\daGPAPIPerAPIAbsMean}{$0.0143 \pm 0.0066$}
\newcommand{\daGPAPIPerAPIAbsStdevNominal}{0.0155}
\newcommand{\daGPAPIPerAPIAbsStdev}{$0.0155 \pm 0.0112$}
\newcommand{\daGPAPIPerAPIParametersTable}{\begin{table} \centering \begin{tabular}{c|S|S} API Version & ${t_0}$ & $\mathrm{decay}$ \\ \hline2	&	0.0	&	0.0237\\3	&	0.0	&	0.0235\\4	&	0.0	&	0.00689\\5	&	36.3	&	0.00429\\6	&	0.0	&	0.0043\\7	&	0.0	&	0.00499\\8	&	47.7	&	0.00415\\9	&	155	&	0.00365\\10	&	90.3	&	0.00364\\11	&	445	&	0.00236\\12	&	356	&	0.00229\\13	&	305	&	0.00237\\14	&	219	&	0.0024\\15	&	161	&	0.00241\\16	&	143	&	0.00229\\17	&	292	&	0.00199\\18	&	136	&	0.00152\\19	&	123	&	0.00158\\\end{tabular} \caption{Parameters for different API versions} \label{tab:per_api_parameters} \end{table}}
\newcommand{\daGPAPISeventeenLaterDate}{2015-09-07}
\newcommand{\daGPAPISeventeenLaterProportion}{79.9\%}
\newcommand{\daGPAPISeventeenEarlierProportion}{20.1\%}
\newcommand{\daAPISeventeenReleaseDate}{2012-10-29}
\newcommand{\daGPAPISeventeenFullDeployment}{2017-09-28}
\newcommand{\daGPAPISeventeenCurveFitParamFirst}{292}
\newcommand{\daGPAPISeventeenCurveFitParamSecond}{\num{0.00199}}
\newcommand{\daGPAPISeventeenCurveFitRMSE}{0.027}
\newcommand{\daGPAPISeventeenCurvePolyRMSE}{0.0103}
\newcommand{\daGPAPISeventeenCurveSplineRMSE}{0.0103}
\newcommand{\daGPAPISeventeenCurveHalfDeployedYearsNominal}{1.75}
\newcommand{\daGPAPISeventeenCurveHalfDeployedYears}{$1.75 \pm 0.07$}
\newcommand{\daGPAPISeventeenCurveHalfDeployedNominal}{639}
\newcommand{\daGPAPISeventeenCurveHalfDeployed}{$639 \pm 27$}
\newcommand{\daGPAPISeventeenCurveFullDeployedYearsNominal}{4.92}
\newcommand{\daGPAPISeventeenCurveFullDeployedYears}{$4.92 \pm 1.07$}
\newcommand{\daGPAPISeventeenCurveFullDeployedNominal}{1\,800}
\newcommand{\daGPAPISeventeenCurveFullDeployed}{$1\,800 \pm 390$}
\newcommand{\daNumUpdatesUpgrades}{3\,690}
\newcommand{\daUpdatesPerMonthPerVersionNominal}{2.59}
\newcommand{\daUpdatesPerMonthPerVersion}{$2.59 \pm 0.04$}
\newcommand{\daNumUpdatesSkippedBig}{3}
\newcommand{\daNumUpdatesBigUpgrades}{843}
\newcommand{\daNumUpdatesDowngrades}{146}
\newcommand{\daNumUpdateFullOnly}{721}
\newcommand{\daPercBigUpgradesNominal}{18.5\%}
\newcommand{\daPercBigUpgrades}{$18.5 \pm 0.6\%$}
\newcommand{\daPercUpdatesDowngradesNominal}{3.2\%}
\newcommand{\daPercUpdatesDowngrades}{$3.2 \pm 0.2\%$}
\newcommand{\daNumDataPoints}{103 billion}
\newcommand{\daMonthsDevices}{2\,140}
\newcommand{\daMonths}{6}
\newcommand{\daAdbEnabledPercNominal}{19.2\%}
\newcommand{\daAdbEnabledPerc}{$19.2 \pm 0.0\%$}
\newcommand{\daNumStartedApps}{129\,000}
\newcommand{\daNumInstalledApps}{184\,000}
\newcommand{\daAPISeventeenReleaseDateMonth}{October 2012}
\newcommand{\daGPAPISeventeenFullDeploymentMonth}{September 2017}
\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.509}
\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.0645}
\newcommand{\daModelHalfDeploymentYears}{0.0822}
\newcommand{\daModelHalfDeployment}{30}
\newcommand{\daModelFullDeploymentYears}{0.888}
\newcommand{\daModelFullDeployment}{324}
\newcommand{\daMeanOutstandingProbability}{1.0\%}
\newcommand{\daMeanOutstandingUncertaintyEquation}{$0.99^n$}
\newcommand{\daGPAPISeventeenLaterDateMonth}{September 2015}
\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & nexus \\ $\pm\sigma$ & 0.211 & 0.169 & 0.175 & 0.632 \\ \hline $u$ & 0.309 & 0.8 & 0.609 & 1.0 \\$m$ & 0.794 & 0.589 & 0.965 & -1.0 \\$f$ & 0.842 & 0.754 & 0.943 & 1.0 \\weight $m$ & 0.976 & 0.968 & 0.987 & 1.0 \\weight $u$ & 0.976 & 1.0 & 0.991 & 1.0 \\equal & 1.0 & 1.0 & 0.996 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics. The uncertainty is constant for each column but does not take into account the uncertainty in the score which produced the ranking.} \label{tab:spearman_ranks} \end{table}}
\newcommand{\daStartDateMonth}{July 2011}
\newcommand{\daEndDateMonth}{September 2015}
\newcommand{\daSecScoreBestsummary}{nexus}
\newcommand{\daSecScoreBestsummaryScoreNominal}{5.28}
\newcommand{\daSecScoreBestsummaryScore}{$5.28 \pm 0.0$}
\newcommand{\daSecScoreBestsummaryNumFullVersions}{99}
\newcommand{\daSecScoreWorstsummary}{non-nexus}
\newcommand{\daSecScoreWorstsummaryScoreNominal}{2.76}
\newcommand{\daSecScoreWorstsummaryScore}{$2.76 \pm 0.0$}
\newcommand{\daSecScoreWorstsummaryNumFullVersions}{1\,280}