### 2016-04-11 avostats.tex
./update.sh

\newcommand{\avoNumSubmitters}{11}
\newcommand{\avoTotalExternalLines}{25 million}
\newcommand{\avoNumExternalProjects}{176}
\newcommand{\avoNumBigExternalLinesOfCode}{25 million}
\newcommand{\avoBigExternalLinesOfCodePerc}{99.7\%}
\newcommand{\avoNumVulnerabilities}{42}
\newcommand{\avoNumVulnAllAndroid}{20}
\newcommand{\avoNumVulnSpecific}{22}
\newcommand{\avoStartDate}{2013-08-28}
\newcommand{\avoEndDate}{2016-03-21}
\newcommand{\avoFirstDataDateMonth}{July 2010}
\newcommand{\avoFirstDataDate}{2010-07-13}
\newcommand{\avoLastDataDateMonth}{October 2015}
\newcommand{\avoLastDataDate}{2015-10-22}
\newcommand{\avoVulnsPerYearNominal}{7.96}
\newcommand{\avoVulnsPerYear}{$7.96 \pm 1.23$}
\newcommand{\avoVulnsPerYearAllAndroidNominal}{3.79}
\newcommand{\avoVulnsPerYearAllAndroid}{$3.79 \pm 0.84$}
\newcommand{\avoVulnsPerYearTwosfNominal}{8.0}
\newcommand{\avoVulnsPerYearTwosf}{$8.0 \pm 1.2$}
\newcommand{\avoVulnsPerYearAllAndroidTwosfNominal}{3.8}
\newcommand{\avoVulnsPerYearAllAndroidTwosf}{$3.8 \pm 0.0$}
\newcommand{\avoTabAndVulns}{\begin{table} \centering \begin{tabular}{l|l|c|c|p{19.5em}} Vulnerability & How known & Date & Categories & CVEs\\ \hline \href{http://androidvulnerabilities.org/vulnerabilities/KillingInTheNameOf_psneuter_ashmem}{KillingInTheNameOf} & Fixed on & 2010-07-13 & system, kernel & CVE-2011-1149~\cite{CVE-2011-1149}\\ \href{http://androidvulnerabilities.org/vulnerabilities/exploid_udev}{exploid udev} & Discovered on & 2010-07-15 & kernel & CVE-2009-1185~\cite{CVE-2009-1185}\\ \href{http://androidvulnerabilities.org/vulnerabilities/levitator}{levitator} & Discovered on & 2011-03-10 & kernel & CVE-2011-1350~\cite{CVE-2011-1350}, CVE-2011-1352~\cite{CVE-2011-1352}\\ \href{http://androidvulnerabilities.org/vulnerabilities/Gingerbreak}{Gingerbreak} & Fixed on & 2011-04-18 & system & CVE-2011-1823~\cite{CVE-2011-1823}\\ \href{http://androidvulnerabilities.org/vulnerabilities/zergRush}{zergRush} & Discovered on & 2011-10-06 & system & CVE-2011-3874~\cite{CVE-2011-3874}\\ \href{http://androidvulnerabilities.org/vulnerabilities/APK_duplicate_file}{APK duplicate file} & Discovered on & 2013-02-18 & signature & ANDROID-8219321, CVE-2013-4787~\cite{CVE-2013-4787}\\ \href{http://androidvulnerabilities.org/vulnerabilities/APK_unchecked_name}{APK unchecked name} & Discovered on & 2013-06-30 & signature & ANDROID-9950697\\ \href{http://androidvulnerabilities.org/vulnerabilities/APK_unsigned_shorts}{APK unsigned shorts} & Fixed on & 2013-07-03 & signature & ANDROID-9695860\\ \href{http://androidvulnerabilities.org/vulnerabilities/Fake_ID}{Fake ID} & Fixed on & 2014-04-17 & signature & \\ \href{http://androidvulnerabilities.org/vulnerabilities/TowelRoot}{TowelRoot} & Discovered on & 2014-05-03 & kernel & CVE-2014-3153~\cite{CVE-2014-3153}\\ \href{http://androidvulnerabilities.org/vulnerabilities/ObjectInputStream_deserializable}{ObjectInputStream} & Discovered on & 2014-06-22 & system & CVE-2014-7911~\cite{CVE-2014-7911}\\ \href{http://androidvulnerabilities.org/vulnerabilities/Stagefright}{Stagefright} & Fixed on & 2015-04-08 & system, network & CVE-2015-1538~\cite{CVE-2015-1538}, CVE-2015-1539~\cite{CVE-2015-1539}, CVE-2015-3824~\cite{CVE-2015-3824}, CVE-2015-3826~\cite{CVE-2015-3826}, CVE-2015-3827~\cite{CVE-2015-3827}, CVE-2015-3828~\cite{CVE-2015-3828}, CVE-2015-3829~\cite{CVE-2015-3829}\\ \href{http://androidvulnerabilities.org/vulnerabilities/One_class_to_rule_them_all}{One class to rule} & Discovered on & 2015-05-22 & system & CVE-2015-3837~\cite{CVE-2015-3837}, CVE-2015-3825~\cite{CVE-2015-3825}, ANDROID-21437603, ANDROID-21583849\\ \href{http://androidvulnerabilities.org/vulnerabilities/Stagefright2}{Stagefright2} & Discovered on & 2015-08-15 & system, network & CVE-2015-6602~\cite{CVE-2015-6602}, CVE-2015-3876~\cite{CVE-2015-3876}\\\end{tabular} \caption{Critical vulnerabilities in Android} \label{tab:andvulns} \end{table}}
\newcommand{\avoNumBigExternalProjects}{40}
\newcommand{\avoNumAnalysedExternalProjects}{28}
\newcommand{\avoNumAnalysedExternalLinesOfCode}{6 million}
\newcommand{\avoAnalysedExternalLinesOfCodePerc}{24.9\%}
\newcommand{\avoBigExternalMedianVersions}{2.0}
\newcommand{\avoBigExternalMeanVersionsNominal}{2.57}
\newcommand{\avoBigExternalMeanVersions}{$2.57 \pm 1.84$}
\newcommand{\avoBigExternalTotalVersions}{72}