### 2016-06-07 dastats.tex
Updated figures etc

\newcommand{\daNumSigModels}{20}
\newcommand{\daSigNumDevices}{100}
\newcommand{\daSigNumDeviceDays}{10\,000}
\newcommand{\daSigNumDays}{100}
\newcommand{\daSigNumDevicesDay}{20}
\newcommand{\daSigVersionPerc}{1\%}
\newcommand{\daSigVersionDays}{10}
\newcommand{\daNumVulnsUsed}{14}
\newcommand{\daNumDeviceDataDevices}{50}
\newcommand{\daOSYearsOfData}{4}
\newcommand{\daOSMonthsOfData}{57}
\newcommand{\daStartDate}{2011-07-01}
\newcommand{\daEndDate}{2016-03-16}
\newcommand{\daFullDeployedAt}{95\%}
\newcommand{\daNumManufacturers}{402}
\newcommand{\daNumSigManufacturers}{11}
\newcommand{\daNumModels}{3\,720}
\newcommand{\daNumOperators}{1\,650}
\newcommand{\daNumSigOperators}{18}
\newcommand{\daNumOSDevices}{24\,600}
\newcommand{\daOSVersionPercValidLines}{99.9\%}
\newcommand{\daOSTotalDaysData}{1\,520\,000}
\newcommand{\daMeanInsecurityPercNominal}{87.6\%}
\newcommand{\daMeanInsecurityPerc}{$87.6 \pm 0.0\%$}
\newcommand{\daMeanInsecurityPercTwosfNominal}{88\%}
\newcommand{\daMeanInsecurityPercTwosf}{$88 \pm 0\%$}
\newcommand{\daMeanSecurityPercNominal}{12.4\%}
\newcommand{\daMeanSecurityPerc}{$12.4 \pm 0.0\%$}
\newcommand{\daMeanSecurityPercTwosfNominal}{12\%}
\newcommand{\daMeanSecurityPercTwosf}{$12 \pm 0\%$}
\newcommand{\daVulnFreeNominal}{0.124}
\newcommand{\daVulnFree}{$0.124 \pm 0.0$}
\newcommand{\daMeanOutstandingVulnerabilities}{$0.661 \pm 0.0$}
\newcommand{\daUpdatednessNominal}{0.0567}
\newcommand{\daUpdatedness}{$0.0567 \pm 0.0$}
\newcommand{\daUpdatednessPercNominal}{5.67\%}
\newcommand{\daUpdatednessPerc}{$5.67 \pm 0.0\%$}
\newcommand{\daUpdatednessPercTwosfNominal}{5.7\%}
\newcommand{\daUpdatednessPercTwosf}{$5.7 \pm 0.0\%$}
\newcommand{\daSecurityScoreNominal}{2.71}
\newcommand{\daSecurityScore}{$2.71 \pm 0.0$}
\newcommand{\daCurrentMeanInsecurityPercNominal}{85.2\%}
\newcommand{\daCurrentMeanInsecurityPerc}{$85.2 \pm 0.0\%$}
\newcommand{\daCurrentMeanInsecurityPercTwosfNominal}{85\%}
\newcommand{\daCurrentMeanInsecurityPercTwosf}{$85 \pm 0\%$}
\newcommand{\daCurrentMeanSecurityPercNominal}{14.8\%}
\newcommand{\daCurrentMeanSecurityPerc}{$14.8 \pm 0.1\%$}
\newcommand{\daCurrentMeanSecurityPercTwosfNominal}{15\%}
\newcommand{\daCurrentMeanSecurityPercTwosf}{$15 \pm 0\%$}
\newcommand{\daCurrentVulnFreeNominal}{0.148}
\newcommand{\daCurrentVulnFree}{$0.148 \pm 0.001$}
\newcommand{\daCurrentMeanOutstandingVulnerabilities}{$0.397 \pm 0.0$}
\newcommand{\daCurrentUpdatednessNominal}{0.0951}
\newcommand{\daCurrentUpdatedness}{$0.0951 \pm 0.0$}
\newcommand{\daCurrentUpdatednessPercNominal}{9.51\%}
\newcommand{\daCurrentUpdatednessPerc}{$9.51 \pm 0.0\%$}
\newcommand{\daCurrentUpdatednessPercTwosfNominal}{9.5\%}
\newcommand{\daCurrentUpdatednessPercTwosf}{$9.5 \pm 0.0\%$}
\newcommand{\daCurrentSecurityScoreNominal}{3.29}
\newcommand{\daCurrentSecurityScore}{$3.29 \pm 0.0$}
\newcommand{\daDailyParticipationNominal}{884}
\newcommand{\daDailyParticipation}{$884 \pm 341$}
\newcommand{\daNumFullVersions}{1\,660}
\newcommand{\daNumSigFullVersions}{117}
\newcommand{\daNumOSVersions}{55}
\newcommand{\daNumSigOSVersions}{30}
\newcommand{\daNumAPIVersions}{17}
\newcommand{\daNumFullVersionUpdates}{5\,970}
\newcommand{\daNumOSVersionUpdates}{4\,310}
\newcommand{\daUpdatesPerYearNominal}{1.43}
\newcommand{\daUpdatesPerYear}{$1.43 \pm 0.01$}
\newcommand{\daUpdatesPerYearTwosfNominal}{1.4}
\newcommand{\daUpdatesPerYearTwosf}{$1.4 \pm 0.0$}
\newcommand{\daNumFullOnlyVersionUpdates}{1\,650}
\newcommand{\daNumSecurityUpdates}{1\,460}
\newcommand{\daNumPossibleSecurityUpdates}{2\,860}
\newcommand{\daOSCurveFitParamFirst}{80.9}
\newcommand{\daOSCurveFitParamSecond}{\num{0.00277}}
\newcommand{\daOSCurveFitRMSE}{0.109}
\newcommand{\daOSCurvePolyRMSE}{0.109}
\newcommand{\daOSCurveSplineRMSE}{0.109}
\newcommand{\daOSCurveHalfDeployedYearsNominal}{0.908}
\newcommand{\daOSCurveHalfDeployedYears}{$0.908 \pm 0.242$}
\newcommand{\daOSCurveHalfDeployedNominal}{332}
\newcommand{\daOSCurveHalfDeployed}{$332 \pm 88$}
\newcommand{\daOSCurveFullDeployedYearsNominal}{3.19}
\newcommand{\daOSCurveFullDeployedYears}{$3.19 \pm 73$}
\newcommand{\daOSCurveFullDeployedNominal}{1\,160}
\newcommand{\daOSCurveFullDeployed}{$1\,160 \pm 268\,000$}
\newcommand{\daAPICurveFitParamFirst}{96.9}
\newcommand{\daAPICurveFitParamSecond}{\num{0.0028}}
\newcommand{\daAPICurveFitRMSE}{0.106}
\newcommand{\daAPICurvePolyRMSE}{0.107}
\newcommand{\daAPICurveSplineRMSE}{0.107}
\newcommand{\daAPICurveHalfDeployedYearsNominal}{0.943}
\newcommand{\daAPICurveHalfDeployedYears}{$0.943 \pm 0.232$}
\newcommand{\daAPICurveHalfDeployedNominal}{344}
\newcommand{\daAPICurveHalfDeployed}{$344 \pm 84$}
\newcommand{\daAPICurveFullDeployedYearsNominal}{3.19}
\newcommand{\daAPICurveFullDeployedYears}{$3.19 \pm 72$}
\newcommand{\daAPICurveFullDeployedNominal}{1\,170}
\newcommand{\daAPICurveFullDeployed}{$1\,170 \pm 264\,000$}
\newcommand{\daVulnAPKDuplicateFileOctoberPerc}{92.2\%}
\newcommand{\daVulnZergRushMonthsDefFixDeployed}{27.6}
\newcommand{\daTabSecScoresmanufacturer}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & $0.34 \pm 0.00$ & $0.33 \pm 0.00$ & $0.74 \pm 0.01$ & $4.28 \pm 0.02$ \\  Motorola & $0.26 \pm 0.00$ & $0.14 \pm 0.00$ & $0.65 \pm 0.01$ & $3.50 \pm 0.02$ \\  HTC & $0.13 \pm 0.00$ & $0.09 \pm 0.00$ & $0.87 \pm 0.01$ & $2.59 \pm 0.02$ \\  Sony & $0.13 \pm 0.00$ & $0.18 \pm 0.00$ & $1.09 \pm 0.02$ & $2.57 \pm 0.02$ \\  Asus & $0.23 \pm 0.00$ & $0.46 \pm 0.01$ & $5.61 \pm 0.06$ & $2.29 \pm 0.02$ \\  Samsung & $0.11 \pm 0.00$ & $0.06 \pm 0.00$ & $0.99 \pm 0.00$ & $2.24 \pm 0.00$ \\  \textit{other} & $0.04 \pm 0.00$ & $0.05 \pm 0.00$ & $1.14 \pm 0.01$ & $1.79 \pm 0.02$ \\  oneplus & $0.02 \pm 0.00$ & $0.31 \pm 0.01$ & $7.85 \pm 0.13$ & $1.00 \pm 0.02$ \\  alps & $0.02 \pm 0.00$ & $0.20 \pm 0.01$ & $5.00 \pm 0.10$ & $0.73 \pm 0.02$ \\  Symphony & $0.00 \pm 0.00$ & $0.10 \pm 0.00$ & $5.74 \pm 0.06$ & $0.32 \pm 0.01$ \\  walton & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $6.08 \pm 0.08$ & $0.28 \pm 0.01$ \\ \end{tabular} \caption{Security scores for manufacturers} \label{tab:sec_manufacturer} \end{table*}}
\newcommand{\daSecScoreBestmanufacturer}{LG}
\newcommand{\daSecScoreBestmanufacturerScoreNominal}{4.28}
\newcommand{\daSecScoreBestmanufacturerScore}{$4.28 \pm 0.0$}
\newcommand{\daSecScoreBestmanufacturerNumFullVersions}{242}
\newcommand{\daSecScoreWorstmanufacturer}{walton}
\newcommand{\daSecScoreWorstmanufacturerScoreNominal}{0.284}
\newcommand{\daSecScoreWorstmanufacturerScore}{$0.284 \pm 0.008$}
\newcommand{\daSecScoreWorstmanufacturerNumFullVersions}{22}
\newcommand{\daTabSecScoresmodel}{\begin{table*} \centering \begin{tabular}{l|l|c|c|c|c} Made by & Model name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline LG & Nexus 5 & $0.71 \pm 0.00$ & $0.65 \pm 0.01$ & $6.09 \pm 0.09$ & $4.80 \pm 0.03$ \\  Samsung & Galaxy Nexus & $0.51 \pm 0.00$ & $0.53 \pm 0.01$ & $1.50 \pm 0.04$ & $4.74 \pm 0.05$ \\  Asus & Nexus 7 & $0.34 \pm 0.00$ & $0.67 \pm 0.01$ & $5.69 \pm 0.08$ & $3.39 \pm 0.03$ \\  LG & Nexus 4 & $0.36 \pm 0.00$ & $0.57 \pm 0.01$ & $5.52 \pm 0.08$ & $3.21 \pm 0.03$ \\  HTC & HTC Desire & $0.14 \pm 0.00$ & $0.07 \pm 0.01$ & $0.51 \pm 0.03$ & $3.03 \pm 0.05$ \\  HTC & Desire HD & $0.09 \pm 0.00$ & $0.05 \pm 0.00$ & $0.43 \pm 0.03$ & $2.86 \pm 0.04$ \\  \textit{unknown} & \textit{other} & $0.08 \pm 0.00$ & $0.13 \pm 0.00$ & $0.74 \pm 0.00$ & $2.65 \pm 0.00$ \\  Samsung & GT-I9000 & $0.03 \pm 0.00$ & $0.03 \pm 0.00$ & $0.44 \pm 0.04$ & $2.55 \pm 0.05$ \\  HTC & HTC Sensation & $0.34 \pm 0.00$ & $0.01 \pm 0.01$ & $1.49 \pm 0.06$ & $2.49 \pm 0.06$ \\  Motorola & DROIDX & $0.02 \pm 0.00$ & $0.04 \pm 0.01$ & $0.55 \pm 0.04$ & $2.39 \pm 0.06$ \\  Samsung & GT-I9100 & $0.21 \pm 0.00$ & $0.01 \pm 0.00$ & $1.26 \pm 0.02$ & $2.23 \pm 0.02$ \\  HTC & HTC Desire S & $0.02 \pm 0.00$ & $0.02 \pm 0.00$ & $1.00 \pm 0.08$ & $1.75 \pm 0.09$ \\  HTC & HTC One & $0.10 \pm 0.00$ & $0.38 \pm 0.01$ & $6.08 \pm 0.12$ & $1.56 \pm 0.04$ \\  Samsung & GT-N7000 & $0.25 \pm 0.00$ & $0.00 \pm 0.00$ & $2.46 \pm 0.05$ & $1.46 \pm 0.03$ \\  Samsung & GT-P1000 & $0.01 \pm 0.00$ & $0.00 \pm 0.01$ & $1.87 \pm 0.06$ & $0.84 \pm 0.05$ \\  Samsung & GT-I9505 & $0.07 \pm 0.00$ & $0.13 \pm 0.00$ & $7.03 \pm 0.07$ & $0.70 \pm 0.01$ \\  Samsung & SM-N9005 & $0.05 \pm 0.00$ & $0.15 \pm 0.01$ & $8.65 \pm 0.20$ & $0.65 \pm 0.03$ \\  Samsung & GT-I9300 & $0.14 \pm 0.00$ & $0.02 \pm 0.00$ & $6.46 \pm 0.05$ & $0.62 \pm 0.01$ \\  HTC & HTC Desire HD & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $3.03 \pm 0.05$ & $0.28 \pm 0.03$ \\  Samsung & GT-N7100 & $0.05 \pm 0.00$ & $0.00 \pm 0.01$ & $6.64 \pm 0.09$ & $0.21 \pm 0.02$ \\  Symphony & Symphony W68 & $0.00 \pm 0.00$ & $0.00 \pm 0.01$ & $11.00 \pm 0.13$ & $0.00 \pm 0.03$ \\ \end{tabular} \caption{Security scores for models} \label{tab:sec_model} \end{table*}}
\newcommand{\daSecScoreBestmodel}{Nexus 5}
\newcommand{\daSecScoreBestmodelScoreNominal}{4.8}
\newcommand{\daSecScoreBestmodelScore}{$4.8 \pm 0.0$}
\newcommand{\daSecScoreBestmodelNumFullVersions}{47}
\newcommand{\daSecScoreWorstmodel}{Symphony W68}
\newcommand{\daSecScoreWorstmodelScoreNominal}{0.0001}
\newcommand{\daSecScoreWorstmodelScore}{$0.0001 \pm 0.0283$}
\newcommand{\daSecScoreWorstmodelNumFullVersions}{1}
\newcommand{\daTabSecScoressummary}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline Nexus devices & $0.53 \pm 0.00$ & $0.50 \pm 0.00$ & $0.69 \pm 0.01$ & $5.63 \pm 0.02$ \\  non-Nexus & $0.09 \pm 0.00$ & $0.02 \pm 0.00$ & $0.74 \pm 0.00$ & $2.35 \pm 0.00$ \\ \end{tabular} \caption{Security scores for Nexus} \label{tab:sec_summary} \end{table*}}
\newcommand{\daTabSecScoresoperator}{\begin{table*} \centering \begin{tabular}{l|c|c|c|c} Name & $f$ & $u$ & $m$ & \textbf{score} \\ &&&& (out of 10) \\ \hline O2 uk & $0.27 \pm 0.00$ & $0.12 \pm 0.00$ & $0.30 \pm 0.01$ & $3.97 \pm 0.02$ \\  Sprint & $0.19 \pm 0.00$ & $0.12 \pm 0.00$ & $0.30 \pm 0.01$ & $3.67 \pm 0.02$ \\  T-Mobile & $0.21 \pm 0.00$ & $0.17 \pm 0.00$ & $0.56 \pm 0.02$ & $3.56 \pm 0.03$ \\  3 & $0.16 \pm 0.00$ & $0.11 \pm 0.00$ & $0.36 \pm 0.02$ & $3.44 \pm 0.03$ \\  Orange & $0.20 \pm 0.00$ & $0.11 \pm 0.00$ & $0.53 \pm 0.03$ & $3.35 \pm 0.04$ \\  AT\&T & $0.12 \pm 0.00$ & $0.09 \pm 0.00$ & $0.37 \pm 0.01$ & $3.20 \pm 0.02$ \\  Vodafone uk & $0.11 \pm 0.00$ & $0.14 \pm 0.00$ & $0.54 \pm 0.03$ & $3.08 \pm 0.04$ \\  Verizon & $0.16 \pm 0.00$ & $0.08 \pm 0.00$ & $0.58 \pm 0.01$ & $3.04 \pm 0.02$ \\  \textit{unknown} & $0.12 \pm 0.00$ & $0.19 \pm 0.00$ & $1.01 \pm 0.02$ & $2.66 \pm 0.02$ \\  n Telenor & $0.06 \pm 0.00$ & $0.13 \pm 0.00$ & $1.31 \pm 0.02$ & $1.93 \pm 0.02$ \\  Airtel & $0.05 \pm 0.00$ & $0.04 \pm 0.00$ & $2.48 \pm 0.04$ & $0.76 \pm 0.02$ \\  Robi & $0.00 \pm 0.00$ & $0.09 \pm 0.00$ & $3.14 \pm 0.05$ & $0.51 \pm 0.02$ \\  Grameenphone & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $3.08 \pm 0.03$ & $0.40 \pm 0.01$ \\  banglalink & $0.00 \pm 0.00$ & $0.04 \pm 0.00$ & $3.12 \pm 0.05$ & $0.37 \pm 0.01$ \\ \end{tabular} \caption{Security scores for operators} \label{tab:sec_operator} \end{table*}}
\newcommand{\daSecScoreBestoperator}{O2 uk}
\newcommand{\daSecScoreBestoperatorScoreNominal}{3.97}
\newcommand{\daSecScoreBestoperatorScore}{$3.97 \pm 0.0$}
\newcommand{\daSecScoreBestoperatorNumFullVersions}{124}
\newcommand{\daSecScoreWorstoperator}{banglalink}
\newcommand{\daSecScoreWorstoperatorScoreNominal}{0.366}
\newcommand{\daSecScoreWorstoperatorScore}{$0.366 \pm 0.013$}
\newcommand{\daSecScoreWorstoperatorNumFullVersions}{76}
\newcommand{\daTabDLDistances}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & Nexus \\ \hline equal & 0.0909 & 0.143 & 0.0 & 0.0 \\weight $u$ & 0.0909 & 0.143 & 0.0 & 0.0 \\weight $m$ & 0.182 & 0.381 & 0.143 & 0.0 \\$f$ & 0.273 & 0.762 & 0.571 & 0.0 \\$m$ & 0.636 & 0.857 & 0.429 & 0.0 \\$u$ & 0.909 & 0.762 & 0.857 & 0.0 \\\end{tabular} \caption{Normalised Damerau-Levenshtein distances for different metrics} \label{tab:dl_distances} \end{table}}
\newcommand{\daTabChangeInScores}{\begin{table*}\small \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & Nexus \\ \hline $m$ & $-1.21 \pm 2.08$ & $-0.819 \pm 2.93$ & $-3.03 \pm 1.84$ & $-2.58 \pm 2.2$ \\weight $m$ & $-0.201 \pm 0.249$ & $-0.116 \pm 0.412$ & $-0.426 \pm 0.25$ & $-0.347 \pm 0.314$ \\equal & $-0.0894 \pm 0.0513$ & $-0.0377 \pm 0.147$ & $-0.137 \pm 0.077$ & $-0.0991 \pm 0.105$ \\weight $u$ & $-0.0671 \pm 0.128$ & $0.00267 \pm 0.17$ & $0.0162 \pm 0.0816$ & $0.0494 \pm 0.0373$ \\$u$ & $0.134 \pm 1.58$ & $0.366 \pm 1.78$ & $1.39 \pm 1.11$ & $1.39 \pm 1.08$ \\$f$ & $0.804 \pm 0.462$ & $0.34 \pm 1.21$ & $1.23 \pm 0.61$ & $0.892 \pm 0.84$ \\\end{tabular} \caption{Mean change in scores for different metrics} \label{tab:change_in_scores} \end{table*}}
\newcommand{\daGPAPICurveFitParamFirst}{83.6}
\newcommand{\daGPAPICurveFitParamSecond}{\num{0.00249}}
\newcommand{\daGPAPICurveFitRMSE}{0.167}
\newcommand{\daGPAPICurvePolyRMSE}{0.167}
\newcommand{\daGPAPICurveSplineRMSE}{0.167}
\newcommand{\daGPAPICurveHalfDeployedYearsNominal}{0.991}
\newcommand{\daGPAPICurveHalfDeployedYears}{$0.991 \pm 0.448$}
\newcommand{\daGPAPICurveHalfDeployedNominal}{362}
\newcommand{\daGPAPICurveHalfDeployed}{$362 \pm 164$}
\newcommand{\daGPAPICurveFullDeployedYearsNominal}{3.52}
\newcommand{\daGPAPICurveFullDeployedYears}{$3.52 \pm 81$}
\newcommand{\daGPAPICurveFullDeployedNominal}{1\,290}
\newcommand{\daGPAPICurveFullDeployed}{$1\,290 \pm 297\,000$}
\newcommand{\daGPAPIPerAPIMeanNominal}{0.00465}
\newcommand{\daGPAPIPerAPIMean}{$0.00465 \pm 0.00294$}
\newcommand{\daGPAPIPerAPIStdevNominal}{0.0197}
\newcommand{\daGPAPIPerAPIStdev}{$0.0197 \pm 0.0118$}
\newcommand{\daGPAPIPerAPIAbsMeanNominal}{0.0135}
\newcommand{\daGPAPIPerAPIAbsMean}{$0.0135 \pm 0.0061$}
\newcommand{\daGPAPIPerAPIAbsStdevNominal}{0.0146}
\newcommand{\daGPAPIPerAPIAbsStdev}{$0.0146 \pm 0.011$}
\newcommand{\daGPAPIPerAPIParametersTable}{\begin{table} \centering \begin{tabular}{c|S|S} API Version & ${t_0}$ & $\mathrm{decay}$ \\ \hline2	&	0.0	&	0.0237\\3	&	0.0	&	0.0235\\4	&	0.0	&	0.00689\\5	&	36.3	&	0.00429\\6	&	0.0	&	0.0043\\7	&	0.0	&	0.00499\\8	&	0.0	&	0.00346\\9	&	155	&	0.00365\\10	&	90.3	&	0.00364\\11	&	446	&	0.00237\\12	&	369	&	0.00237\\13	&	306	&	0.00238\\14	&	220	&	0.00241\\15	&	162	&	0.00242\\16	&	153	&	0.00236\\17	&	296	&	0.00205\\18	&	146	&	0.00162\\19	&	137	&	0.0017\\21	&	103	&	0.00121\\22	&	127	&	0.00102\\\end{tabular} \caption{Parameters for different API versions} \label{tab:per_api_parameters} \end{table}}
\newcommand{\daGPAPISeventeenLaterDate}{2016-04-04}
\newcommand{\daGPAPISeventeenLaterProportion}{87.3\%}
\newcommand{\daGPAPISeventeenEarlierProportion}{12.7\%}
\newcommand{\daAPISeventeenReleaseDate}{2012-10-29}
\newcommand{\daGPAPISeventeenFullDeployment}{2017-08-23}
\newcommand{\daGPAPISeventeenCurveFitParamFirst}{296}
\newcommand{\daGPAPISeventeenCurveFitParamSecond}{\num{0.00205}}
\newcommand{\daGPAPISeventeenCurveFitRMSE}{0.0258}
\newcommand{\daGPAPISeventeenCurvePolyRMSE}{0.0182}
\newcommand{\daGPAPISeventeenCurveSplineRMSE}{0.0182}
\newcommand{\daGPAPISeventeenCurveHalfDeployedYearsNominal}{1.74}
\newcommand{\daGPAPISeventeenCurveHalfDeployedYears}{$1.74 \pm 0.07$}
\newcommand{\daGPAPISeventeenCurveHalfDeployedNominal}{634}
\newcommand{\daGPAPISeventeenCurveHalfDeployed}{$634 \pm 25$}
\newcommand{\daGPAPISeventeenCurveFullDeployedYearsNominal}{4.82}
\newcommand{\daGPAPISeventeenCurveFullDeployedYears}{$4.82 \pm 0.97$}
\newcommand{\daGPAPISeventeenCurveFullDeployedNominal}{1\,760}
\newcommand{\daGPAPISeventeenCurveFullDeployed}{$1\,760 \pm 355$}
\newcommand{\daNumUpdatesUpgrades}{4\,080}
\newcommand{\daUpdatesPerMonthPerVersionNominal}{2.38}
\newcommand{\daUpdatesPerMonthPerVersion}{$2.38 \pm 0.03$}
\newcommand{\daNumUpdatesSkippedBig}{3}
\newcommand{\daNumUpdatesBigUpgrades}{891}
\newcommand{\daNumUpdatesDowngrades}{146}
\newcommand{\daNumUpdateFullOnly}{1\,640}
\newcommand{\daPercBigUpgradesNominal}{15.2\%}
\newcommand{\daPercBigUpgrades}{$15.2 \pm 0.5\%$}
\newcommand{\daPercUpdatesDowngradesNominal}{2.49\%}
\newcommand{\daPercUpdatesDowngrades}{$2.49 \pm 0.20\%$}
\newcommand{\daNumDataPoints}{44 billion}
\newcommand{\daMonthsDevices}{2\,110}
\newcommand{\daMonths}{6}
\newcommand{\daAdbEnabledPercNominal}{19.4\%}
\newcommand{\daAdbEnabledPerc}{$19.4 \pm 0.0\%$}
\newcommand{\daNumStartedApps}{134\,000}
\newcommand{\daNumInstalledApps}{234\,000}
\newcommand{\daAPISeventeenReleaseDateMonth}{October 2012}
\newcommand{\daGPAPISeventeenFullDeploymentMonth}{August 2017}
\newcommand{\daMeanOutstandingVulnerabilitiesNominal}{0.661}
\newcommand{\daCurrentMeanOutstandingVulnerabilitiesNominal}{0.397}
\newcommand{\daModelHalfDeploymentYears}{0.0384}
\newcommand{\daModelHalfDeployment}{14}
\newcommand{\daModelFullDeploymentYears}{0.69}
\newcommand{\daModelFullDeployment}{252}
\newcommand{\daMeanOutstandingProbability}{1.0\%}
\newcommand{\daMeanOutstandingUncertaintyEquation}{$0.99^n$}
\newcommand{\daGPAPISeventeenLaterDateMonth}{April 2016}
\newcommand{\daTabSpearmanRanks}{\begin{table} \centering \begin{tabular}{l|c|c|c|c} & manufacturer & model & operator & Nexus \\ $\pm\sigma$ & 0.2 & 0.141 & 0.175 & 0.632 \\ \hline $u$ & 0.3 & 0.712 & 0.552 & 1.0 \\$m$ & 0.855 & 0.517 & 0.934 & 1.0 \\$f$ & 0.964 & 0.694 & 0.943 & 1.0 \\weight $m$ & 0.973 & 0.948 & 0.991 & 1.0 \\equal & 0.991 & 0.996 & 1.0 & 1.0 \\weight $u$ & 0.991 & 0.996 & 1.0 & 1.0 \\\end{tabular} \caption{Spearman Rank correlation coefficients for different metrics. The uncertainty is constant for each column but does not take into account the uncertainty in the score which produced the ranking.} \label{tab:spearman_ranks} \end{table}}
\newcommand{\daStartDateMonth}{July 2011}
\newcommand{\daEndDateMonth}{March 2016}
\newcommand{\daSecScoreBestsummary}{Nexus devices}
\newcommand{\daSecScoreBestsummaryScoreNominal}{5.63}
\newcommand{\daSecScoreBestsummaryScore}{$5.63 \pm 0.0$}
\newcommand{\daSecScoreBestsummaryNumFullVersions}{150}
\newcommand{\daSecScoreWorstsummary}{non-Nexus devices}
\newcommand{\daSecScoreWorstsummaryScoreNominal}{2.35}
\newcommand{\daSecScoreWorstsummaryScore}{$2.35 \pm 0.0$}
\newcommand{\daSecScoreWorstsummaryNumFullVersions}{1\,630}
\newcommand{\daMeanDefInsecurityPercNominal}{73.2\%}
\newcommand{\daMeanDefInsecurityPerc}{$73.2 \pm 0.0\%$}
\newcommand{\daMeanDefInsecurityPercTwosfNominal}{73\%}
\newcommand{\daMeanDefInsecurityPercTwosf}{$73 \pm 0\%$}
\newcommand{\daMeanMaybeSecurePercNominal}{16.4\%}
\newcommand{\daMeanMaybeSecurePerc}{$16.4 \pm 0.0\%$}
\newcommand{\daMeanMaybeSecurePercTwosfNominal}{16\%}
\newcommand{\daMeanMaybeSecurePercTwosf}{$16 \pm 0\%$}
\newcommand{\daOptSecurityScoreNominal}{3.29}
\newcommand{\daOptSecurityScore}{$3.29 \pm 0.0$}
\newcommand{\daCurrentMeanDefInsecurityPercNominal}{69.4\%}
\newcommand{\daCurrentMeanDefInsecurityPerc}{$69.4 \pm 0.0\%$}
\newcommand{\daCurrentMeanDefInsecurityPercTwosfNominal}{69\%}
\newcommand{\daCurrentMeanDefInsecurityPercTwosf}{$69 \pm 0\%$}
\newcommand{\daCurrentMeanMaybeSecurePercNominal}{19.1\%}
\newcommand{\daCurrentMeanMaybeSecurePerc}{$19.1 \pm 0.1\%$}
\newcommand{\daCurrentMeanMaybeSecurePercTwosfNominal}{19\%}
\newcommand{\daCurrentMeanMaybeSecurePercTwosf}{$19 \pm 0\%$}
\newcommand{\daCurrentOptSecurityScoreNominal}{3.92}
\newcommand{\daCurrentOptSecurityScore}{$3.92 \pm 0.0$}